Knutson
```{r}
require(terra)
require(tidyverse)
require(pgirmess)
require(ncf)
library(geoR)
```


# Background

Researchers set up a square plot with a dimension of 130 m on each side. Within the plot they established blue vane traps in a 5 m grid pattern to sample bee abundance. Every day for 10 days the researchers emptied the blue vane traps and recorded the number of bees collected. After the 10-day sampling period, they calculated the mean daily number of bees captured in the trap each day.

# Challenge 1 (2 points)

Import the blueVanes.csv file from GitHub. Plot the mean bee abundance as a function of the x and y coordinates. Upon initial inspection, does there appear to be spatial dependence in the bee response data?

```{r}
thebees=read.csv("https://raw.githubusercontent.com/ValenteJJ/SpatialEcology/main/Week6/blueVanes.csv")
thebees$coordinates=paste(thebees$x, thebees$y, sep = "+")
head(thebees)
# Plot thebees as a function of coordinates
ggplot(thebees, aes(x=x, y=y, color=meanBee))+
  geom_point()+
  theme_bw()+
  theme(panel.grid=element_blank())+
  scale_color_gradient(low='white', high='purple')
```
There appears to be a bit of positive spatial dependence in the high and low mean abundance. lower mean bee abundance looks like it occurs more often in the bottom left corner, top right corner and upper left central. Higher mean bee abundances appear more often in the bottom right and left central 

$\color{red}{\text{Excellent. +2}}$



______________________________________________________________________________________________________________
# Challenge 2 (3 points)

Empirically test for evidence of spatial dependence in bee abundance by first building a corelogram, then building a semi-variogram. Be sure to provide an envelope for assessing whether the values in each figure are statistically different from what would be expected if the bee values were randomly distributed. Describe the patterns you see, and your interpretation with regards to the spatial dependence in mean bee abundance. Be sure to identify the maximum distance at which you should interpret your figures (2/3 of the maximum distance between points).

```{r}
#CORRELOGRAM
coords = cbind(thebees$x, thebees$y)#create matrix of distances among all the points
colnames(coords) = c('x', 'y')
distMatrix = as.matrix(dist(coords))
maxDistance = (2/3) * max(distMatrix) #define the appropriate spatial distance

correlogPgirmessBees = pgirmess::correlog(coords, thebees$meanBee, method='Moran', nbclass=14, alternative = 'two.sided')

round(correlogPgirmessBees, 2)

tmpBees = data.frame(correlogPgirmessBees) %>%
  mutate(significant = ifelse(p.value < 0.05, 'yes', 'no'))

ggplot(tmpBees, aes(x=dist.class, y=coef))+
  geom_line()+
  geom_point(aes(color=significant, size=n))+
  scale_color_manual(values=c('black', 'red'))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  geom_hline(yintercept=0, linetype='dashed')+
  geom_vline(xintercept=maxDistance, color='red')+
  scale_x_continuous(breaks=seq(0, 75, 5))


#VARIOGRAM
geobees = as.geodata(thebees)
empGeorbees = variog(geobees, max.dist=maxDistance, breaks=seq(0, maxDistance, by=3))
empEnvbees = variog.mc.env(geobees, obj.var=empGeorbees, nsim=99)
plot(empGeorbees, envelope=empEnvbees)
```

in the correleogram, we see evidence of positive 0-20 meters, negative spatial dependence between 25-50 meters and after 80 meters, and then evidence of positive spatial dependence after 80 meters. In the variogram.

$\color{red}{\text{In the variogram, you're actually seeing evidence that variance among points > 80 m apart is greater than would be expected by chance, indicating negative spatial dependence. +2.5}}$



_____________________________________________________________________________________________________________________________________________________________________
# Challenge 3 (4 points)



Empirically test for evidence of spatial autocorrelation IN THE MODEL RESIDUALS by first building a correlogram, then building a semi-variogram. Be sure to provide an envelope for assessing whether the values in each figure are statistically different from what would be expected if the bee values were randomly distributed in space. Compare these figures to those built for the raw data. What similarities or differences to you notice?
```{r}
#LINEAR MODEL: Fit a simple linear model (assuming a normally distributed error term) that models mean bee abundance as a function of an intercept only (no other covariates in the model). Compare the model-estimated intercept to the bee abundance values averaged across all sites. What do you notice?  
beelm = lm(meanBee ~ 1, data=thebees)  # fitted linear model in which 1 signifies an intercept-only model
summary(beelm)
plot(beelm)
intercept = coef(beelm)[1]# pull out the intercept
avgBeeMean = mean(thebees$meanBee)# average bee mean across all sites

# Compare the model-estimated intercept to the mean abundance, this is just aesthetic-   ease of interpretation listing them next to each other. heavy use of ChatGPT here
print(paste("Model-estimated intercept:", round(intercept, 2)))
print(paste("Mean abundance across all sites:", round(avgBeeMean, 2)))


#____________________________________________________



#spatial autocorrelation IN THE MODEL RESIDUALS by building a correlogram and semi-variogram. Be sure to provide an envelope for assessing whether the values in each figure are statistically different from what would be expected if the bee values were randomly distributed in space. Compare these figures to those built for the raw data. What similarities or differences to you notice?
 
 #EXTRACT RESIDUALS: 
  BeeResidual = residuals(beelm)
  
#CORRELOGRAM
# Create new data frame with linear model residuals in place of bee abundance column
thebeeslm = thebees
thebeeslm = thebees[,-3]
thebeeslm$BeeResidual = BeeResidual

# Create Correlogram
correlogPgirmess_lmresid = pgirmess::correlog(coords, thebeeslm$BeeResidual, method='Moran', nbclass=26, alternative = 'two.sided') 
round(correlogPgirmess_lmresid, 2)

# Plot Correlogram
tmpbeelm = data.frame(correlogPgirmess_lmresid) %>%
  mutate(significant = ifelse(p.value < 0.05, 'yes', 'no')) # <-- In Place of Envelope

ggplot(tmpbeelm, aes(x=dist.class, y=coef))+
  geom_line()+
  geom_point(aes(color=significant, size=n))+
  scale_color_manual(values=c('black', 'red'))+
  theme_bw()+
  theme(panel.grid=element_blank())+
  geom_hline(yintercept=0, linetype='dashed')+
  
  scale_x_continuous(breaks=seq(0, 130, 10)) +
    labs(title = "Correlogram: Simple Linear Model Residuals for bees",
       x = "Distance (m)",
       y = "Coefficent")

#VARIOGRAM
geobeeslm = as.geodata(thebeeslm, data.col=4)
empGeorbeeslm = variog(geobeeslm, max.dist=maxDistance, breaks=seq(0, maxDistance, by=3))
empEnvbeeslm = variog.mc.env(geobeeslm, obj.var=empGeorbeeslm, nsim=99)
plot(empGeorbeeslm, envelope=empEnvbeeslm)
```
the intercept and average bee abundance appear to be the same (13.93). this is because there are no other independent variables in this model, and the intercept is representing the mean abundance. this suggests that the lm is not seeing variability in the abundance other than what is represented by the bee mean average.

$\color{red}{\text{And what about the similarities/differences between the correlogram and variogram produced here vs. the one produced in the previous challenge? +3}}$


# Challenge 4 (4 points)

Conventional wisdom indicates that the distribution of floral resources may substantially influence bee abundance. In addition, I am  completely making up the following fact: the average bee has a perceptual range of about 15 m. Below I have provided code for importing a raster that shows the number of flowers counted in 1 m grid cells across the entire study area. Use the focal() function from the previous lab to generate a new raster in which each cell value represents the mean number of flowers in cells within a 15 m radius. Plot the distribution of the blue vane traps over the 15 m focal raster. No text required here, just show me the plot.

```{r}

floral = rast('https://github.com/ValenteJJ/SpatialEcology/raw/main/Week6/floralResources.tif')
focal_window=focalMat(floral,15,type='circle', fillNA = FALSE)
# Perform the focal operation
r_focal = focal(floral, w = focal_window, fun = 'sum')

# Plot the original and focal raster
plot(floral, main = "Original Raster")
plot(r_focal, main = "Focal Raster")
FlowerDist = as.data.frame(r_focal, xy = TRUE)

# Plot with ggplot2
ggplot() +
  geom_raster(data = FlowerDist, aes(x = x, y = y, fill = focal_sum)) +
  geom_point(data = thebees, aes(x = x, y = y, color = meanBee)) +
  scale_fill_viridis_c() + 
  scale_color_gradient(low = "white", high = "purple", name = "Mean Bee Abundance") +
  labs(title = "Flower Distribution and Mean Bee Abundance",
       x = "X Coordinate", y = "Y Coordinate") +
  theme_minimal()
```

$\color{red}{\text{Good. +4}}$


# Challenge 5 (4 points)

Extract the focal flower value from the raster you just created at each point using the extract() function. You now have an estimate for the average number of flowers per meter around each blue vane trap. Fit a second linear model, but this time include average floral abundance within 15 m as a covariate. Export the residuals from this fitted model, and again, examine a corelogram and variogram of the residuals. As always, be sure to plot random envelopes. Now what do you notice?

```{r}
    
#CORRELOGRAM
  # Setting Up Model
    coords = as.matrix(thebees[,1:2])
    thebees$flowerRes = terra::extract(r_focal, coords)[,1]
    FloralLm = lm(meanBee ~ flowerRes, data = thebees)#floral linear model
    FloralLm_residuals = residuals(FloralLm)#extract her residuals
    
   # Create Correlogram
    FloralCorrelLm = pgirmess::correlog(coords, FloralLm_residuals , method='Moran', nbclass=26, alternative = 'two.sided')
    round(FloralCorrelLm, 2)
    tmpflowerlm = data.frame(FloralCorrelLm) %>%
      mutate(significant = ifelse(p.value < 0.05, 'yes', 'no')) # <-- In Place of Envelope
    
    ggplot(tmpflowerlm, aes(x=dist.class, y=coef))+
      geom_line()+
      geom_point(aes(color=significant, size=n))+
      scale_color_manual(values=c('black', 'red'))+
      theme_bw()+
      theme(panel.grid=element_blank())+
      geom_hline(yintercept=0, linetype='dashed')+
      
      scale_x_continuous(breaks=seq(0, 130, 10)) +
        labs(title = "Correlogram: Simple Linear Model Residuals for flowers",
           x = "Distance (m)",
           y = "Coefficent")


#VARIOGRAM
flowerslm = thebees
flowerslm = thebees[-4]
flowerslm$FloralLm_residuals = FloralLm_residuals
geoflowerlm = as.geodata(flowerslm,data.col=5)
empGeorflowerlm = variog(geoflowerlm, max.dist=maxDistance, breaks=seq(0, maxDistance, by=3))
empEnvflowerlm = variog.mc.env(geoflowerlm, obj.var=empGeorflowerlm, nsim=99)
plot(empGeorflowerlm, envelope=empEnvflowerlm)
```
the variogram and the correlogram of the flower residuals indicate that the data falls within what would be expected if random.

$\color{red}{\text{Be clear with your words here. We're talking about if the residuals were randomly distributed in space. +4}}$



____________________________________________________________________________________________
# Challenge 6 (3 points)

What is your assessment regarding the distribution of bee abundance in this study area? Does it exhibit spatial dependence, and if so, what is driving that pattern?

It seems as if there is some spatial dependence of the bees within about 20 meters; however, this is explained by the flower distribution.

$\color{red}{\text{Specifically, there is spatial aggregation in the flowers, the bees are responding to the flowers, so there is spatial aggregation in the bees. But, if we model the effect of flowers, then the residual abundance is randomly distributed. +2.5}}$

